---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Brazilian Real Estate

<!-- badges: start -->
<!-- badges: end -->

The goal of realestatebr is to facilitate the access to reports and indicators on the Brazilian real estate market. This package currently covers only the residential market but in the future it will also encompass other real estate markets.

**Important**: This package is still under development but can already be installed. Feedback is welcome.

## Installation

You can install the development version of realestatebr from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("viniciusoike/realestatebr")
```

## Downloading data

To get the data use one of the `get_*` functions such as `get_abecip_indicators`. 

```{r example}
library(realestatebr)
# Import data from Abecip
abecip <- get_abecip_indicators()

abecip
```

Outputs will be either a named `list` of `tibbles` or a single `tibble`. All `get_*` functions have two arguments: `category` and `cached`. The `category` argument helps to filter the final result and will always default to `'all'`. The `cached` argument is a `logical` that indicates whether to download the data directly from this GitHub repository. This option will usually be faster.

```{r}
sbpe <- get_abecip_indicators(category = "sbpe")

sbpe
```

## Residential Property Price Indexes

There are several house price indices available in the Brazilian residential real estate market. The `get_rppi_*` functions collect all of these indices. A general `get_rppi()` function

```{r}
# For better plots
library(ggplot2)

# Download and clean all sales RPPIs
rppi <- get_rppi(category = "sale", stack = TRUE)
# Filter only Brazil
rppi_brazil <- subset(rppi, name_muni == "Brazil" & date >= as.Date("2019-01-01"))

ggplot(rppi_brazil, aes(date, acum12m)) +
  geom_line(aes(color = source)) +
  scale_y_continuous(labels = scales::label_percent()) +
  theme_light()
```

International comparisons are also possible using the BIS data

```{r}
library(dplyr, warn.conflicts = FALSE)
# Download simplified BIS RPPI data
bis <- get_rppi_bis()
# Highlight some countries, show only real indices
bis_brasil <- bis |> 
  filter(
    country %in% c("Australia", "Brazil", "Chile", "Japan", "United States"),
    is_nominal == FALSE,
    date >= as.Date("2000-01-01")
    )

ggplot(bis_brasil, aes(date, index)) +
  geom_line(aes(color = country)) +
  geom_hline(yintercept = 100) +
  theme_light() +
  theme(legend.position = "top")
```
