---
title: "Getting Started with realestatebr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with realestatebr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = FALSE
)
```

```{r setup, message = FALSE}
library(realestatebr)
library(dplyr)
library(ggplot2)
```

`realestatebr` provides a unified interface to Brazilian real estate data from
multiple public sources. All datasets are returned as tidy `tibble` objects.

## Core Interface

Three functions cover most use cases.

**`list_datasets()`** returns a catalogue of all available datasets and their
tables.

```{r}
list_datasets()
```

**`get_dataset()`** retrieves any dataset by name. Without a `table` argument
it returns the default table; use `table` to select a specific sub-table.

```{r}
#| eval: false
# Default table
abecip <- get_dataset("abecip")

# Specific table
sbpe <- get_dataset("abecip", table = "sbpe")
```

**`get_dataset_info()`** shows available tables and metadata for a given
dataset.

```{r}
#| eval: false
info <- get_dataset_info("abecip")
names(info$categories)
#> [1] "sbpe"  "units"  "cgi"
```

## The `source` Argument

The `source` argument controls where data comes from. The default (`"auto"`)
checks the local cache first, then falls back to the GitHub release. Typically, the best option is to use the default or `"github"`.

```{r}
#| eval: false
get_dataset("abecip", source = "cache")    # local cache (instant, works offline)
get_dataset("abecip", source = "github")   # GitHub release
get_dataset("abecip", source = "fresh")    # direct from the original source
```

Cache files are stored in the user data directory and can be inspected with
`list_cached_files()` or cleared with `clear_user_cache()`.

## Example: Housing Credit Cycle

SBPE (Sistema Brasileiro de Poupança e Empréstimo) is the primary funding
vehicle for residential mortgages in Brazil. The table tracks monthly inflows,
outflows, and the resulting net credit flow.

```{r}
sbpe <- get_dataset("abecip", table = "sbpe")

# Annual net credit flow
sbpe_annual <- sbpe |>
  filter(date >= as.Date("2019-01-01")) |>
  mutate(year = lubridate::year(date)) |>
  summarise(net_flow = sum(sbpe_netflow, na.rm = TRUE) / 1e3, .by = year)

ggplot(sbpe_annual, aes(year, net_flow)) +
  geom_col(fill = "steelblue", alpha = 0.9) +
  labs(
    title = "SBPE: Annual Net Housing Credit",
    x     = NULL,
    y     = "R$ billions"
  ) +
  theme_minimal()
```

The companion table `"units"` contains monthly counts of financed units broken
down by programme (SBPE, FGTS, CGI). Joining the two tables gives a
credit-per-unit metric:

```{r}
units <- get_dataset("abecip", table = "units")

# SBPE units financed per year
units_annual <- units |>
  filter(date >= as.Date("2019-01-01")) |>
  mutate(year = lubridate::year(date)) |>
  summarise(sbpe_units = sum(units_total, na.rm = TRUE), .by = year)

# Implied credit per unit (R$ thousands)
sbpe_annual |>
  left_join(units_annual, by = "year") |>
  mutate(credit_per_unit = net_flow * 1e6 / sbpe_units) |>
  ggplot(aes(year, credit_per_unit)) +
  geom_line(color = "steelblue", linewidth = 0.9) +
  labs(
    title = "SBPE: Average Credit per Financed Unit",
    x     = NULL,
    y     = "R$ thousands"
  ) +
  theme_minimal()
```

## Example 2: Real Estate Credit Portfolio

The `bcb_realestate` dataset imports all real estate statistics from the Brazilian Central Bank. This is a relatively large dataset and exploring can be cumbersome. Each series is uniquely identified dy `date` and `series_info`. Helper functions `v1`, `v2`, ..., `v5`, `abbrev_state`, `category`, and `type` are provided to simplify exploring the dataset.

```{r}
bcb <- get_dataset("bcb_realestate")

# Get a specific series
sfh_pf <- bcb |>
  filter(series_info == "credito_estoque_carteira_credito_pf_sfh_br")

# Get all series for a specific category
# All credit lines for brazilian households
credit_stock <- bcb |>
  filter(
    category == "credito",
    type == "estoque",
    v1 == "carteira",
    v2 == "credito",
    v3 == "pf",
    v5 == "br"
  )

ggplot(credit_stock, aes(date, value)) +
  geom_area(aes(fill = v4), alpha = 0.9)
```

As a final warning, note that the `bcb_realestate` dataset follows the `YYYY-MM-DD` format with the last day of the month (e.g. `2023-01-31`). This can cause issues when merging with other datasets, since the first day of the month is the more common date format. To avoid this, use `lubridate::floor_date(date, 'month')`. Future versions of `realestatebr` might provide this as a default behavior.

## Next Steps

- `vignette("working-with-rppi")` — property price indices in depth
- `?get_dataset` — full parameter reference
