---
title: "Getting Started with realestatebr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with realestatebr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Installation

```{r, eval = FALSE}
# Install from GitHub
remotes::install_github("viniciusoike/realestatebr")
```

## Quick Start

The realestatebr package provides a unified interface for accessing Brazilian real estate data from multiple sources.

```{r setup, message=FALSE, eval=FALSE}
library(realestatebr)
library(dplyr)
library(ggplot2)
```

## Basic Usage

### Discovering Datasets

Use `list_datasets()` to see all available data:

```{r, eval=FALSE}
# List all available datasets
datasets <- list_datasets()
head(datasets)

# Filter by source
bcb_data <- list_datasets(source = "BCB")
```

### Getting Data

All data is accessed through `get_dataset()`:

```{r, eval=FALSE}
# Get dataset (returns default table)
abecip <- get_dataset("abecip")
head(abecip)

# Get specific table
sbpe <- get_dataset("abecip", table = "sbpe")

# See available tables for a dataset
info <- get_dataset_info("abecip")
names(info$categories)
```

### Data Sources

Control where data comes from with the `source` parameter:

```{r, eval=FALSE}
# Auto (default): Try GitHub cache first, fallback to fresh download
data <- get_dataset("secovi", source = "auto")

# GitHub cache only (faster, may be outdated)
cached <- get_dataset("secovi", source = "github")

# Fresh download (slower, always current)
fresh <- get_dataset("secovi", source = "fresh")
```

## Available Datasets

The package currently provides access to these datasets:

**Credit & Finance:**
- **abecip**: Housing credit data (SBPE flows, units, home equity)
- **bcb_realestate**: Real estate credit and market data from BCB

**Market Indicators:**
- **abrainc**: Primary market indicators (launches, sales, business conditions)
- **secovi**: São Paulo market data (condos, rentals, launches, sales)

**Price Indices:**
- **rppi**: Brazilian property price indices (FipeZap, IVGR, IGMI, IVAR, IQA, Secovi-SP)
- **rppi_bis**: International property price indices from BIS (60+ countries)

**Economic Data:**
- **bcb_series**: Economic time series from BCB
- **cbic**: Cement consumption and production data

**Other:**
- **fgv_ibre**: FGV economic indicators

Use `list_datasets()` to see the full list with details.

## Example Workflows

### Example 1: Housing Credit Analysis

```{r, eval=FALSE}
# Get SBPE housing credit data
sbpe <- get_dataset("abecip", table = "sbpe")

# Plot net flow over time
ggplot(sbpe, aes(x = date, y = netflow)) +
  geom_line(color = "steelblue") +
  labs(title = "SBPE Net Flow",
       x = NULL,
       y = "R$ (millions)") +
  theme_minimal()
```

### Example 2: Real Estate Credit by State

```{r, eval=FALSE}
# Get BCB real estate data
bcb <- get_dataset("bcb_realestate")

# Aggregate credit by date
credit_data <- bcb |>
  filter(category == "credit", type == "stock") |>
  summarise(total = sum(value, na.rm = TRUE), .by = date)

# Plot
ggplot(credit_data, aes(x = date, y = total)) +
  geom_line(color = "steelblue") +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Total Real Estate Credit Stock",
       x = NULL,
       y = "R$ (billions)") +
  theme_minimal()
```

### Example 3: São Paulo Market Data

```{r, eval=FALSE}
# Get Secovi São Paulo data
secovi <- get_dataset("secovi")

# See available categories
info <- get_dataset_info("secovi")
names(info$categories)

# Get specific categories
condo_fees <- get_dataset("secovi", table = "condo")
rental_data <- get_dataset("secovi", table = "rent")
```

## Working with Multi-Table Datasets

Some datasets have multiple tables (categories). Use `get_dataset_info()` to explore:

```{r, eval=FALSE}
# Get dataset structure
info <- get_dataset_info("cbic")

# See available tables
names(info$categories)

# Access specific tables
cement_monthly <- get_dataset("cbic", table = "cement_monthly_consumption")
cement_prices <- get_dataset("cbic", table = "cement_cub_prices")
```

## Controlling Output

### Verbosity

```{r, eval=FALSE}
# Show progress messages (default)
data <- get_dataset("abecip", quiet = FALSE)

# Suppress messages
data <- get_dataset("abecip", quiet = TRUE)
```

### Error Handling

The package provides informative error messages:

```{r, eval=FALSE}
# Invalid dataset name
get_dataset("invalid_name")
#> Error: Dataset 'invalid_name' not found
#> ℹ Available datasets: abecip, abrainc, bcb_realestate, ...
#> ℹ Use list_datasets() to see all available datasets

# Invalid table
get_dataset("abecip", table = "wrong_table")
#> Error: Table 'wrong_table' not found for dataset 'abecip'
#> ℹ Available tables: sbpe, units, cgi
```

## Migration from v0.3.x

**Breaking Change in v0.4.0**: Individual `get_*()` functions have been removed.

```{r, eval=FALSE}
# OLD (v0.3.x) - No longer works
# abecip_old <- get_abecip_indicators(table = "sbpe")

# NEW (v0.4.0) - Use get_dataset()
abecip_new <- get_dataset("abecip", table = "sbpe")
```

See [NEWS.md](../NEWS.md) for complete migration details.

## Next Steps

- See `vignette("working-with-rppi")` for detailed property price index guide
- Use `?get_dataset` for full documentation
- Visit the [package website](https://viniciusoike.github.io/realestatebr/) for more examples
