---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

```{r setup}
library(realestatebr)
# Additional packages to facilitate data visualization.
library(ggplot2)
library(dplyr)
library(tidyr)
library(RcppRoll)
```

# General Usage

The `realestatebr` package aims to offer real estate reports and data in a "tidy" output. By "tidy" I mean a form that is more amenable to be used in data analysis work. Currently, the package is focused only on residential real estate but in the future the package may also encompass commercial real estate information.

This package downloads, cleans and returns data. All functions follow a similar syntax (`get_*()`) with two arguments:

-   `category` - A string that indicates specific parts/tables of the data (will default to `'all'`)
-   `cached` - A logical that indicates if the data should be downloaded from the original provider or from the GitHub repository.

Both arguments have the same purpose: to accelerate data consumption. By specifying a particular table, the function can save processing time; similarly, downloading the cleaned data from the GitHub repository can be faster.

One of the exceptions to the rule above is the `get_rppi()` function, which downloads data on Residential Property Prices Indexes. It has an additional third argument `stack` which is a logical that indicates whether the indexes should be returned in a single table or as elements in a named `list`.

The specific `get_rppi_*()` functions have different arguments.

# Examples

## Abrainc

As a simple use case example consider the case of downloading the quarterly [Abrainc-Fipe report](https://www.abrainc.org.br/lab). To get all of the available data use the `get_abrainc_indicators()` function.

```{r}
abrainc <- get_abrainc_indicators()
```

The result will be a named list with three `tibble`.

```{r}
abrainc
```

When working specifically with a single table this can be specified in the `category` argument.

```{r}
# Either subset the existing abrainc file or download again
radar <- get_abrainc_indicators(category = "radar")

# Plot example
ggplot(radar, aes(x = date, y = value)) +
  geom_line() +
  facet_wrap(vars(variable), ncol = 3) +
  theme_light()
```

## Brazilian Central Bank

The Brazilian Central Bank has a wide variety of real estate data that is stored into a single long-format csv file. This makes working with this data quite difficult. The example below shows how to find the total amount of new real estate credit in three specific states.

```{r}
bcb <- get_bcb_realestate(cached = TRUE)

credit_south <- bcb %>%
  filter(
    category == "credito",
    type == "contratacao",
    v1 == "contratado",
    v2 == "pf",
    abbrev_state %in% c("PR", "RS", "SC")
  ) %>%
  summarise(total = sum(value, na.rm = TRUE), .by = c("date", "abbrev_state"))

credit_south <- credit_south %>%
  group_by(abbrev_state) %>%
  mutate(
    trend = RcppRoll::roll_mean(total, n = 12, fill = NA),
    trend = RcppRoll::roll_mean(trend, n = 2, fill = NA)
  ) %>%
  ungroup()

ggplot(credit_south, aes(x = date, color = abbrev_state)) + 
  geom_line(aes(y = total), alpha = 0.5) +
  geom_line(aes(y = trend)) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(breaks = seq(0, 3, 0.5) * 1e9, labels = seq(0, 3, 0.5)) +
  scale_color_brewer(name = "State", type = "qual", palette = 6) +
  labs(
    title = "Monthly Real Estate Credit Volume Trend",
    subtitle = "Smoothed 12x2-month moving averge of monthly real estate credit operations.",
    x = NULL,
    y = "R$ (Billion)",
    caption = "Source: Brazilian Central Bank (BCB)"
  ) +
  theme_light() +
  theme(legend.position = "bottom")
```

The `category` argument allows for a more direct "query"; also, data is presented in "wide" format.

```{r}
index <- get_bcb_realestate("indices", cached = TRUE)

index
```


## RPPI

### Brazil

There are several Residential Property Price Indices in Brazil. Each index has its own methodology, coverage, and data source. The table below summarizes

| Name        | Rent/sale     | Coverage                     | Source           | Methodology                   |
|-------------|---------------|------------------------------|------------------|-------------------------------|
| IVG-R       | Sale          | National                     | Bank appraisals  | Median Stratified + Smoothing |
| FipeZap     | Rent and Sale | National (cities)            | Online Listings  | Median Stratified + Smoothing |
| IGMI-R      | Sale          | National (cities)            | Bank appraisals  | Hedonic                       |
| IVAR        | Rent          | National (cities)            | Rental Contracts | Repeat-rent                   |
| QuintoAndar | Rent          | Rio de Janeiro and São Paulo | Rental Contracts | Median Stratified             |

To get a specific series use one of the `get_rppi_*` functions. The IVAR index is available for many cities.

```{r}
ivar <- get_rppi_ivar()

ggplot(ivar, aes(x = date, y = index)) +
  geom_line() +
  facet_wrap(vars(name_muni))
```

The FipeZap Index has many cities and subindices.

```{r}
fipe <- get_rppi_fipezap()

subfipe <- fipe |> 
  filter(
    name_muni == "São Paulo",
    market == "residential",
    variable == "index",
    rooms != "total"
    )

ggplot(subfipe, aes(x = date, y = value, color = rooms)) +
  geom_line() +
  facet_wrap(vars(rent_sale)) +
  theme_light()
```

### International

The Bank for International Settlements compiles RPPIs from across the world. The default output calls the selected quarterly series.

```{r}
bis <- get_bis_rppi()

subbis <- bis |> 
  filter(
    date >= as.Date("2000-01-01"),
    reference_area %in% c("World", "Japan", "United States", "Brazil"),
    is_nominal == TRUE,
    unit == "Index, 2010 = 100"
    )

ggplot(subbis, aes(x = date, y = value, color = reference_area)) +
  geom_line() +
  theme_light()
```


