---
title: "Working with Property Price Indices"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Property Price Indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = FALSE
)
```

```{r setup, message = FALSE}
library(realestatebr)
library(dplyr)
library(ggplot2)

theme_series <- function() {
  theme_minimal(base_family = "Avenir", base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.line.x = element_line(color = "gray10", linewidth = 0.5),
      axis.ticks.x = element_line(color = "gray10", linewidth = 0.5),
      legend.position = "bottom",
      palette.color.discrete = c(
        "#1E3A5F",
        "#DD6B20",
        "#2C7A7B",
        "#D69E2E",
        "#805AD5",
        "#C53030"
      )
    )
}
```

Brazil has several residential property price indices (RPPIs) from different
institutions, each with distinct methodologies, geographic coverage, and time
spans. This vignette shows how to access and work with each one.

All indices are available through `get_dataset("rppi", table = <index>)`.

## Recommended Workflow

### The best indices

The two most consistent indices are the IGMI-R (for sale prices) and the IVG-R (for rent prices). The only available option for commercial prices is FipeZap.

The IGMI-R is a hedonic price index with a robust dataset (transaction prices). The only shortcoming of the index is its relatively short time span (2014-present) and geographic coverage (major cities only). The IVAR is a repeat-rent index with a robust dataset (residential rental contracts). Unfortunately, the dataset has a very short time span (2019-present) and even more limited geographic coverage (only 4 cities).

FipeZap has the broadest geographic (50+ cities) and temporal coverage (2008-present), but is based on median listing prices rather than transaction prices. As a result, the index is much less reliable.

### Working with realestatebr

For easier comparison, the `table` parameter accepts a `"sale"` or `"rent"` argument, which stakcs all available indices into a single `tibble`.

```{r}
sale <- get_dataset("rppi", table = "sale")
```

## Sale Indices

### IVG-R

The IVG-R (Índice de Valores de Garantia de Imóveis Residenciais) is published
by the Central Bank of Brazil. It covers the national market from 2001 and is
widely regarded as the official RPPI in Brazil. The index is based on median
appraisal prices from mortgage contracts, smoothed with a 3-month moving
average and the HP filter.

```{r}
ivgr <- get_dataset("rppi", "ivgr")

ggplot(ivgr, aes(date, index)) +
  geom_line(color = "#27708C", linewidth = 0.9) +
  labs(
    title = "IVG-R — National Sale Index",
    x     = NULL,
    y     = "Index"
  ) +
  theme_series()
```

### IGMI-R

The IGMI-R (Índice Geral do Mercado Imobiliário Residencial) is published by
Abecip/FGV. It uses a hedonic pricing model applied to actual transaction
prices, making it the most methodologically rigorous city-level index
available. Coverage spans major Brazilian cities from 2014.

```{r}
igmi <- get_dataset("rppi", "igmi")

igmi |>
  filter(
    name_muni %in%
      c("São Paulo", "Rio De Janeiro", "Belo Horizonte", "Brasília")
  ) |>
  ggplot(aes(date, index, color = name_muni)) +
  geom_line(linewidth = 0.8) +
  labs(
    title = "IGMI-R — Sale Index by City",
    x = NULL,
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

### FipeZap (sale)

FipeZap is the broadest RPPI in terms of geographic coverage, with indices for
more than 50 cities. It is based on median **listing prices** rather than
transaction prices. Available from 2008 for both residential and commercial
markets, with breakdowns by number of bedrooms. Prices are stratified by region and income level, using Census weights, and the final series is smoothed using a 3-month moving average.

Key columns specific to FipeZap: `market` (residential/commercial),
`rent_sale`, `rooms` (1, 2, 3, 4+, total), `variable` (index, chg, acum12m,
price_m2, yield).

```{r}
fz <- get_dataset("rppi", "fipezap")

subzap <- fz |>
  filter(
    market == "residential",
    rent_sale == "sale",
    rooms == "total",
    variable == "acum12m",
    name_muni %in% c("São Paulo", "Rio De Janeiro", "Belo Horizonte")
  )

ggplot(aes(date, value, color = name_muni)) +
  geom_line(linewidth = 0.8) +
  labs(
    title = "FipeZap — Residential Sale Index",
    x = NULL,
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

## Rent Indices

### IVAR

The IVAR (Índice de Variação de Aluguéis Residenciais) is published by FGV and
uses a repeat-rent methodology applied to actual rental contracts. It covers
major Brazilian cities from 2019 and is available at both city and national
level. Notably, the index has no smoothing, making it rather volatile.

```{r}
ivar <- get_dataset("rppi", "ivar")

# National index with 6-month moving average
ivar |>
  filter(name_muni == "Brazil") |>
  mutate(ma6 = zoo::rollmean(index, 6, fill = NA, align = "right")) |>
  ggplot(aes(date)) +
  geom_line(aes(y = index), color = "#A6511F", linewidth = 0.5, alpha = 0.5) +
  geom_line(aes(y = ma6), color = "#27708C", linewidth = 0.9) +
  labs(
    title = "IVAR — National Rent Index",
    subtitle = "Raw (orange) vs 6-month moving average (blue)",
    x = NULL,
    y = "Index"
  ) +
  theme_series()
```

```{r}
# City-level comparison
ivar |>
  filter(name_muni != "Brazil") |>
  ggplot(aes(date, index, color = name_muni)) +
  geom_line(linewidth = 0.8) +
  labs(
    title = "IVAR — Rent Index by City",
    x = NULL,
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

### IQA and IQAIW

QuintoAndar published two successive rental indices for São Paulo and Rio de
Janeiro. The **IQA** (2019–mid-2023) was based on median contract prices. In
mid-2023 it rebranded to **IQAIW** (Índice QuintoAndar ImovelWeb) and adopted
a hedonic model incorporating both listing and contract prices. The two series
are not directly comparable due to the methodology break.

The original IQA compared median contract prices, stratified by region and income level (using Census weights, similar to FipeZap). The IQAIW uses a hedonic model (spatial GAM), double imputation, and a similar Census-based stratification.

Notably, the IQA returns rent prices (R$/m²) rather than an index, so a base-100 normalisation is needed to build an index.

```{r}
iqa <- get_dataset("rppi", "iqa")
iqaiw <- get_dataset("rppi", "iqaiw")

# Normalise IQA to base 100 at first observation
iqa_idx <- iqa |>
  mutate(index = rent_price / first(rent_price) * 100, .by = name_muni)

# Stack and compare in São Paulo
bind_rows(
  mutate(iqa_idx, source = "IQA"),
  mutate(iqaiw, source = "IQAIW")
) |>
  filter(name_muni == "São Paulo") |>
  ggplot(aes(date, index, color = source)) +
  geom_line(linewidth = 0.9) +
  labs(
    title = "QuintoAndar Rent Index — São Paulo",
    subtitle = "IQA (pre-2023) and IQAIW (post-2023) use different methodologies",
    x = NULL,
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

### FipeZap (rent)

FipeZap provides the broadest geographic coverage for rental indices, but is
based on median listing prices rather than contracts. The example below shows
the 12-month accumulated change across cities with complete data from 2021.

```{r}
fz <- get_dataset("rppi", "fipezap")

start_date <- as.Date("2021-01-01")

# Keep only cities present at start_date to avoid gaps
cities_available <- fz |>
  filter(
    market    == "residential",
    rooms     == "total",
    variable  == "acum12m",
    rent_sale == "rent",
    date      == start_date,
    !is.na(value)
  ) |>
  pull(name_muni)

fz |>
  filter(
    market    == "residential",
    rooms     == "total",
    variable  == "acum12m",
    rent_sale == "rent",
    date      >= start_date,
    name_muni %in% cities_available,
    !is.na(value)
  ) |>
  ggplot(aes(date, value)) +
  geom_line(linewidth = 0.7, color = "#27708C") +
  facet_wrap(vars(name_muni)) +
  labs(
    title = "FipeZap — 12-month Rent Change by City",
    x     = NULL,
    y     = "Accumulated 12-month change (%)"
  ) +
  theme_series()
```

## Comparing Indices

### Stacked tables

The `"sale"` and `"rent"` tables stack all available indices into a single
data frame, making cross-source comparisons straightforward.

```{r}
sale_indices <- get_dataset("rppi", "sale")
unique(sale_indices$source)

rent_indices <- get_dataset("rppi", "rent")
unique(rent_indices$source)
```

### Normalising to a common base

Each index has its own base period, so normalisation is required before
comparing levels. The example below rebases all national sale indices to
January 2018 = 100.

```{r}
sales <- get_dataset("rppi", "sale")

national <- sales |>
  filter(name_muni == "Brazil", date >= as.Date("2018-01-01"))

national_rebased <- national |>
  mutate(
    index_rebased = index / first(index) * 100,
    .by = source
  )

# Annotate with total growth
total_growth <- national_rebased |>
  summarise(
    growth    = last(index_rebased) - first(index_rebased),
    date      = last(date),
    index_rebased = last(index_rebased),
    .by = source
  ) |>
  mutate(label = sprintf("%s: +%.1f%%", source, growth))

ggplot(national_rebased, aes(date, index_rebased, color = source)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 100, linetype = "dashed", alpha = 0.4) +
  geom_label(
    data   = total_growth,
    aes(label = label),
    hjust  = 0,
    nudge_x = 30,
    show.legend = FALSE,
    size   = 3
  ) +
  scale_x_date(limits = c(as.Date("2018-01-01"), as.Date("2025-09-01"))) +
  labs(
    title = "Brazil National Sale Indices — Rebased to Jan 2018",
    x     = NULL,
    y     = "Index (Jan 2018 = 100)",
    color = NULL
  ) +
  theme_series()
```

### 12-month growth rates

Growth rates are directly comparable across indices and avoid the rebasing
issue. Note that different methodologies may point to divergent trends.

```{r}
national_rebased |>
  ggplot(aes(date, acum12m, color = source)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.4) +
  scale_y_continuous(labels = scales::label_percent(scale = 100)) +
  labs(
    title = "Brazil National Sale Indices — 12-month Change",
    x     = NULL,
    y     = NULL,
    color = NULL
  ) +
  theme_series()
```

### Rent vs sale divergence

The post-pandemic divergence between rental and sale markets is one of the
most discussed trends in Brazilian real estate. FipeZap, which covers both
markets, makes this comparison straightforward.

```{r}
fz <- get_dataset("rppi", "fipezap")

base_year <- 2019

base_avg <- fz |>
  filter(
    name_muni %in% c("São Paulo", "Rio De Janeiro", "Belo Horizonte"),
    market   == "residential",
    rooms    == "total",
    variable == "index",
    lubridate::year(date) == base_year
  ) |>
  summarise(base = mean(value, na.rm = TRUE), .by = c(name_muni, rent_sale))

fz |>
  filter(
    name_muni %in% c("São Paulo", "Rio De Janeiro", "Belo Horizonte"),
    market   == "residential",
    rooms    == "total",
    variable == "index",
    date     >= as.Date("2019-01-01")
  ) |>
  left_join(base_avg, by = c("name_muni", "rent_sale")) |>
  mutate(idx = 100 * value / base) |>
  ggplot(aes(date, idx, color = rent_sale)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 100, linetype = "dashed", alpha = 0.4) +
  facet_wrap(vars(name_muni)) +
  labs(
    title    = "Post-pandemic Rent vs Sale Divergence",
    subtitle = "FipeZap residential index, 2019 average = 100",
    x        = NULL,
    y        = "Index",
    color    = NULL
  ) +
  theme_series()
```

## International: BIS

The BIS dataset provides quarterly residential property price indices for 60+
countries, enabling Brazil to be placed in international context.

```{r}
bis <- get_dataset("rppi_bis")

bis |>
  filter(
    reference_area %in% c("Brazil", "United States", "Germany", "Japan"),
    is_nominal == FALSE,
    unit       == "Index, 2010 = 100",
    date       >= as.Date("2000-01-01")
  ) |>
  ggplot(aes(date, value, color = reference_area)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 100, linetype = "dashed", alpha = 0.4) +
  labs(
    title    = "Real Residential Property Prices",
    subtitle = "BIS, index 2010 = 100",
    x        = NULL,
    y        = "Index",
    color    = NULL
  ) +
  theme_series()
```

## Index Reference

| Index | Source | Coverage | Methodology | From |
|-------|--------|----------|-------------|------|
| IVG-R | BCB | National | Median appraisal prices, HP-filtered | 2001 |
| IGMI-R | ABECIP/FGV | Major cities | Hedonic (transaction prices) | 2014 |
| FipeZap | FIPE/ZAP | 50 cities | Median listing prices | 2008 |
| IVAR | FGV | Major cities + national | Repeat-rent (contracts) | 2019 |
| IQA | QuintoAndar | São Paulo, Rio | Median contract prices | 2019 |
| IQAIW | QuintoAndar | São Paulo, Rio | Hedonic (listing + contracts) | 2023 |
| Secovi-SP | SECOVI-SP | São Paulo | Median prices (transactions) | 2008 |
