---
title: "Housing Credit in Brazil"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  fig.width = 7,
  fig.asp = 0.618,
  fig.align = "center",
  warning = FALSE,
  message = FALSE
)
```

Brazilian housing finance is largely organized around the SFH (Sistema
Financeiro da Habitação), in which mortgages are funded by savings deposits
(SBPE) and by the FGTS workers' fund. Three datasets cover this market.
`abecip`, `bcb_series`, and `bcb_realestate` appear below in increasing order
of detail.

| Table name | Source | Description |
|----|----|----|
| `abecip` | Abecip | Savings flows and financed units by the SBPE system. |
| `bcb_series` | Brazilian Central Bank (SGS) | Economic time series related to credit and housing credit (e.g. lending volumes, rates, delinquency, etc.). |
| `bcb_realestate` | Brazilian Central Bank (Estatísticas do Mercado Imobiliário) | A more detailed dataset containing economic time series related to housing credit. |

`realestatebr` follows its sources closely, so monetary values come in nominal
Brazilian reais with no inflation adjustment. This holds for all three
datasets used here.

```{r}
library(realestatebr)
library(dplyr)
library(ggplot2)
```

The code below defines a common theme for the plots in this article and can be
omitted.

```{r}
if (requireNamespace("ekioplot", quietly = TRUE)) {
  color_palette <- ekioplot::ekio_pal()
  theme_series <- function() {
    ekioplot::theme_ekio(base_size = 10) +
      theme(
        palette.color.discrete = color_palette
      )
  }
} else {
  color_palette <- c(
    "#1E3A5F",
    "#DD6B20",
    "#2C7A7B",
    "#D69E2E",
    "#805AD5",
    "#C53030"
  )
  theme_series <- function() {
    theme_minimal(base_size = 10) +
      theme(
        legend.position = "bottom",
        palette.color.discrete = color_palette
      )
  }
}
```

## Funding

SBPE savings accounts are an important funding source for mortgages, so
sustained withdrawals eventually constrain new lending. The `sbpe` table from
`abecip` tracks monthly deposits, withdrawals, and the resulting stock.

```{r}
sbpe <- get_dataset("abecip", table = "sbpe")

glimpse(sbpe)
```

The plot shows the accumulated net flow over twelve months. Periods of persistent outflows, such as the years after 2021, signal tighter funding ahead.

```{r}
sbpe_flows <- sbpe |>
  filter(date >= as.Date("2012-01-01")) |>
  mutate(
    netflow_12m = zoo::rollsumr(sbpe_netflow, k = 12, fill = NA) / 1e3
  )

ggplot(sbpe_flows, aes(date, netflow_12m)) +
  geom_area(fill = color_palette[1], alpha = 0.9) +
  geom_hline(yintercept = 0) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "SBPE Net Savings Flow",
    subtitle = "Accumulated net flow in 12 months",
    y = "R$ (billion)"
  ) +
  theme_series()
```

The rest of the funding market shows up in `bcb_realestate`. That dataset takes
some time to learn, since a chain of `v1` to `v5` components identifies each
series, but the query below stays short. It returns total credit funding from
private sources for real estate loans in Brazil.

```{r}
bcb_re <- get_dataset("bcb_realestate")

bcb_funding <- bcb_re |>
  filter(category == "fontes", v1 == "br") |>
  mutate(funding_source = factor(type), value_bln = value / 1e9)
```

```{r}
ggplot(bcb_funding, aes(date, value_bln)) +
  geom_area(aes(fill = funding_source)) +
  scale_fill_manual(values = color_palette, labels = \(x) toupper(x)) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  labs(
    title = "Market Funding Sources",
    subtitle = "Private source credit funding for real estate loans",
    y = "R$ (billion)",
    fill = NULL
  ) +
  theme_series()
```

## Lending

The Central Bank reports the flow of new real estate loans to households. That
series sits in the `core` table of `bcb_series`, which gathers the most
relevant real estate credit series. We filter on `name_simplified` here, but
the dataset also carries `code_bcb`, which matches the original SGS codes.

```{r}
bcb <- get_dataset("bcb_series", table = "core")

fimob <- bcb |>
  filter(name_simplified == "fimob_pf_total") |>
  mutate(
    lending_12m = zoo::rollsumr(value, k = 12, fill = NA) / 1e3
  )
```

As mentioned above, these values are nominal (i.e. not adjusted for inflation).

```{r}
ggplot(filter(fimob, !is.na(lending_12m)), aes(date, lending_12m)) +
  geom_line(color = color_palette[1], lwd = 0.8) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "New Mortgage Lending to Households",
    subtitle = "Earmarked real estate operations, accumulated in 12 months",
    y = "R$ (billion)"
  ) +
  theme_series()
```

In physical terms, the `units` table from `abecip` counts how many homes were financed with SBPE funds, split between construction and acquisition loans.

```{r}
units <- get_dataset("abecip", table = "units")

units_12m <- units |>
  select(date, units_construction, units_acquisition) |>
  tidyr::pivot_longer(-date, names_to = "type", values_to = "units") |>
  mutate(
    units_12m = zoo::rollsumr(units, k = 12, fill = NA) / 1e3,
    type_label = if_else(
      type == "units_construction",
      "Construction",
      "Acquisition"
    ),
    .by = type
  )
```

```{r}
ggplot(filter(units_12m, !is.na(units_12m)), aes(date, units_12m)) +
  geom_line(aes(color = type_label), lwd = 0.8) +
  scale_color_manual(values = color_palette[c(1, 2)]) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "Units Financed with SBPE Funds",
    subtitle = "Accumulated in 12 months",
    y = "Units (thousand)",
    color = NULL
  ) +
  theme_series()
```

## The price of credit

Mortgage rates move with monetary policy but are partly insulated by regulated
pricing within the SFH. The `core` table includes both the average rate charged
on new operations and the average cost of the outstanding stock, which adjusts
much more slowly.

```{r}
rates <- bcb |>
  filter(name_simplified %in% c("taxa_fimob_pf_total", "icc_fimob")) |>
  mutate(
    rate_label = if_else(
      name_simplified == "icc_fimob",
      "Outstanding stock (average cost)",
      "New operations (average rate)"
    )
  )

rates_recent <- filter(rates, date >= as.Date("2012-01-01"))
```

```{r}
ggplot(rates_recent, aes(date, value)) +
  geom_line(aes(color = rate_label), lwd = 0.8) +
  scale_color_manual(values = color_palette[c(1, 2)]) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "Household Mortgage Rates",
    subtitle = "Earmarked real estate credit operations (% per year)",
    y = "%",
    color = NULL
  ) +
  theme_series()
```

For a more detailed view of credit rates, we turn to `bcb_realestate`. The code
below finds the rates for all credit lines at the national (`BR`) level. In
these series `pf` stands for `pessoa física` (private individual) and `pj` for
`pessoa jurídica` (company).

```{r}
credit_rates <- bcb_re |>
  filter(
    category == "credito",
    v1 == "taxa",
    # only for families/individuals (i.e. pessoa física)
    v2 == "pf",
    # weighted average for Brazil
    abbrev_state == "BR"
  ) |>
  mutate(
    credit_label = dplyr::recode(
      v3,
      comercial = "Commercial",
      fgts = "FGTS",
      `home-equity` = "Home Equity",
      livre = "Market",
      sfh = "SFH"
    )
  )
```

The plot shows the average rate applied across each credit line, making the
size of the spread clear.

```{r}
ggplot(credit_rates, aes(date, value)) +
  geom_line(aes(color = credit_label), lwd = 0.8) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    title = "Household Mortgage Rates",
    subtitle = "Detailed view of credit rates (% per year)",
    y = "%",
    color = NULL
  ) +
  theme_series()
```

## Credit risk

Delinquency in housing credit is typically low because loans are
collateralized, although borrowers can still fall behind when economic
conditions deteriorate.

```{r}
risk <- bcb |>
  filter(
    name_simplified %in%
      c(
        "inad_credito_direcionado_pf_fimob",
        "atraso_fimob_pf_total"
      )
  ) |>
  mutate(
    risk_label = if_else(
      name_simplified == "atraso_fimob_pf_total",
      "Arrears (15-90 days)",
      "Past due (90+ days)"
    )
  )
```

```{r}
ggplot(risk, aes(date, value)) +
  geom_line(aes(color = risk_label), lwd = 0.8) +
  scale_color_manual(values = color_palette[c(2, 1)]) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "Household Mortgage Delinquency",
    subtitle = "Share of earmarked real estate credit operations",
    y = "%",
    color = NULL
  ) +
  theme_series()
```

Once again, for a more detailed view we use `bcb_realestate`. The code below
finds the risk classification of the stock of SFH loans.

```{r}
sfh_risk_stock <- bcb_re |>
  filter(
    category == "credito",
    type == "estoque",
    v1 == "risco-operacao",
    # only for families/individuals (i.e. pessoa física)
    v2 == "pf",
    v3 == "sfh",
    v5 == "br"
  ) |>
  mutate(
    credit_risk = factor(
      v4,
      levels = c("aa", "a", "b", "c", "d-mais"),
      labels = c("AA", "A", "B", "C", "D+")
    )
  )
```

```{r}
ggplot(sfh_risk_stock, aes(date, value)) +
  geom_area(aes(fill = credit_risk)) +
  scale_fill_manual(values = color_palette) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  labs(
    title = "Risk Classification of SFH Loans",
    subtitle = "Share of total SFH credit loans classified by risk credit (following resolution 2.682/1999)",
    y = "% (share of stock)",
    fill = NULL
  ) +
  theme_series()
```

## Where the credit is

The `bcb_realestate` dataset has regional data by state. The code below sums the
household credit portfolio across credit lines for each state's latest
available month.

```{r}
stock_states <- bcb_re |>
  filter(
    category == "credito",
    type == "estoque",
    v1 == "carteira",
    v2 == "credito",
    v3 == "pf",
    abbrev_state != "BR"
  ) |>
  filter(date == max(date), .by = abbrev_state) |>
  summarise(stock = sum(value) / 1e9, .by = abbrev_state) |>
  slice_max(stock, n = 10)
```

```{r}
ggplot(stock_states, aes(stock, reorder(abbrev_state, stock))) +
  geom_col(fill = color_palette[1], width = 0.8) +
  scale_x_continuous(expand = expansion(c(0, 0.05))) +
  labs(
    title = "Household Credit Stock by State",
    subtitle = "Ten largest states, latest available observation",
    y = NULL,
    x = "R$ (billion)"
  ) +
  theme_series() +
  theme(
    panel.grid.major.x = element_line(),
    panel.grid.major.y = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_text(),
    axis.line.y = element_line(color = "gray10", linewidth = 0.5),
    axis.ticks.y = element_line(color = "gray10", linewidth = 0.5)
  )
```

## Home equity lending

Home equity loans (CGI, Crédito com Garantia de Imóvel) remain a small but growing market. The `cgi` table from `abecip` summarizes contracts, balances, and default rates.

```{r}
cgi <- get_dataset("abecip", table = "cgi")

glimpse(cgi)
```

```{r}
ggplot(cgi, aes(date, stock_contracts / 1e3)) +
  geom_line(color = color_palette[1], lwd = 0.8) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "Outstanding Home Equity Contracts",
    y = "Contracts (thousand)"
  ) +
  theme_series()
```

## Learn more

The help topics `?abecip`, `?bcb_series`, and `?bcb_realestate` document the
columns of each dataset used here. For the primary market, see [The Primary
Market and the Construction Cycle](primary-market.html); for price indices, see
[Working with Property Price Indices](working-with-rppi.html).
