---
title: "MCMV data dictionary"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

registry_path <- system.file(
  "extdata",
  "datasets.yaml",
  package = "realestatebr"
)
mcmv <- yaml::read_yaml(registry_path)$datasets$mcmv

column_table <- function(table) {
  category <- mcmv$categories[[table]]
  sources <- category$source_columns
  dictionary <- data.frame(
    Column = paste0("`", vapply(sources, `[[`, "", "name"), "`"),
    `Source header` = paste0("`", names(sources), "`"),
    Type = vapply(sources, `[[`, "", "type"),
    Unit = vapply(sources, `[[`, "", "unit"),
    check.names = FALSE
  )
  dictionary$Definition <- unlist(
    category$columns[vapply(sources, `[[`, "", "name")],
    use.names = FALSE
  )
  knitr::kable(dictionary, row.names = FALSE)
}
```

This dictionary defines the three MCMV tables returned by
`query_dataset("mcmv")`. The column tables below come from the package
registry, the same source as the `?mcmv` help page. The
[working article](working-with-mcmv.html) shows how to query the tables.

The source is the Ministério das Cidades'
[MCMV open-data page](`r mcmv$url`). Its official
[data dictionary](https://www.gov.br/cidades/pt-br/acesso-a-informacao/acoes-e-programas/habitacao/programa-minha-casa-minha-vida/minha-casa-minha-vida-fnhis-sub-50-1/arquivos-fnhis-sub-50/Dicionarios_SNH_2025_10_09.pdf)
(version 2.4, October 2025) predates the July 2026 files and does not
describe several of their changes. The notes below record where the published
files depart from it.

## Conventions

- **Names.** Columns use English names. Measures share prefixes: `units_*`
  for housing units, `amount_*` for financed, contracted, or disbursed
  amounts, and `subsidy_<source>_<component>` for subsidy components.
- **Money.** Amounts are nominal BRL, stored as doubles without rounding.
- **Municipalities.** `code_muni_6` is the six-digit IBGE code, without the
  check digit. `name_muni`, `state`, and `region` keep the source's spelling
  and capitalization.
- **Dates.** `reference_date` is the date the source reports for its own
  extract, not the download or publication date. Contract dates come from
  timestamps that are always midnight, so they are stored as dates.
- **Missing values.** `NA` means the source left the value blank, or the
  release did not publish the column at all. The package never fills in
  missing values, from zero or from older snapshots.
- **Categories.** Labels and codes appear as published, including
  inconsistent capitalization. The package applies no crosswalk the source
  does not document.

## `financing`

`r mcmv$categories$financing$description`. In the July 2026 release, every
row reports exactly one unit.

```{r financing}
column_table("financing")
```

### Categorical values

| Column | Values in the July 2026 release | Meaning |
|---|---|---|
| `income_band` | `1`, `2`, `3`, `4`, missing | Undocumented codes. Code `4` appears only in contracts signed from 2025 onward. Older releases published text labels instead of codes. |
| `fgts_account_holder` | `S`, `N`, `1`, `0`, missing | FGTS account holder flag, mixing two encodings. Missing in about 94% of rows. |
| `sex` | `M`, `F`, `X`, missing | Source codes. `X` is undocumented. |
| `property_type` | `novo`/`usado` in three capitalizations, missing | New or existing dwelling. |
| `amortization_system` | `price`, `sac`, `sacre` in varying case, missing | Price, SAC, or SACRE amortization. |
| `financing_program` | Program labels | Includes lines such as `Fundo Social` and `Classe Média`. Some labels vary in spelling, such as `Apoio à Produção` and `Apoio à producao`. |

`interest_rate` is in percentage units as published. The source does not
state its periodicity. `household_income` has no stated periodicity either.

### Subsidy components

The four subsidy columns split the subsidy by funding source and mechanism.

- `subsidy_fgts_discount` and `subsidy_ogu_discount` are the discounts
  applied to the purchase, funded by the FGTS or by the federal budget (OGU).
- `subsidy_fgts_interest` and `subsidy_ogu_interest` are the interest
  subsidies ("equilíbrio") that lower the financing cost.

In the July 2026 release, the four components sum to `subsidy_total` in
`financing_summary` for every municipality, month, and income band. The
discount components alone do not.

## `financing_summary`

`r mcmv$categories$financing_summary$description`. These are the Ministry's
own totals, not sums computed by the package.

```{r summary}
column_table("financing_summary")
```

In the July 2026 release, the summary and `financing` agree in every
municipality, year, month, and income band: unit counts match exactly and
amounts match within BRL 0.01.

## `subsidized_projects`

`r mcmv$categories$subsidized_projects$description`. These projects are
funded by the federal budget (OGU) under modalities such as FAR, Entidades,
Rural, and Oferta Publica.

```{r projects}
column_table("subsidized_projects")
```

Keep these points in mind before aggregating.

- **Duplicates.** The June 2026 release has 3,281 exact duplicate rows.
  Remove them with `distinct()` before counting projects or units.
- **Operation codes.** `operation_code` is missing in 4,608 rows and repeats
  in others, so it is not a primary key.
- **Unit accounting.** `units_contracted` equals `units_delivered +
  units_outstanding + units_cancelled` in every row where all four are
  reported. Six rows lack `units_cancelled`.
- **Labels.** `modality` includes both `Rural` and `RURAL`.

`responsible_entity_cnpj` and `postal_code` are text, which preserves leading
zeros.

## Changes across source releases

The Ministry has changed file layouts between releases without notice. The
package recognizes the layouts below and rejects any other layout.

| Release | Table | Change |
|---|---|---|
| December 2025 | `financing` | 23 columns with birth date and project name; income bands as text labels; Brazilian number format. |
| March 2026 | `financing` | No `sex` column; state header renamed; decimal-point numbers; about 494,000 rows labeled `Fora MCMV/CVA`. |
| July 2026 | `financing` | `sex` returns; birth date and project name removed; income bands become codes `1` to `4`; Brazilian number format. |
| Up to 2025 | `financing_summary` | Annual totals by municipality, without month, region, or income band. |
| July 2026 | `financing_summary` | Monthly totals by municipality and income band; decimal-point numbers. |
| March 2026 | `subsidized_projects` | Decimal-point numbers. |
| June 2026 | `subsidized_projects` | Same headers as March, but Brazilian number format and thousands separators in municipality codes. |

Columns missing from a release appear as `NA` in every row of that snapshot.
In annual summaries, a missing `contract_month` means the source did not
publish monthly detail, not January.

Each release replaces the full history, and releases overlap. They are
published as separate snapshots and should not be appended to each other.
