---
title: "The Primary Market and the Construction Cycle"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618,
  fig.align = "center",
  message = FALSE,
  warning = FALSE
)
```

This article follows the development cycle, from launches to sales and
deliveries, and the business environment around it. Three datasets do the
work.

- `abrainc` for national primary market indicators from the ABRAINC-FIPE
  panel of developers
- `secovi` for the São Paulo market
- `fgv_ibre` for construction costs and business sentiment

The aim is to show what these datasets hold, not to settle any question about
the market. Besides the usual `dplyr` and `ggplot2`, we use the `trendseries`
package to estimate trends and cycles.

```{r}
library(realestatebr)
library(trendseries)
library(dplyr)
library(ggplot2)
```

The code below defines a common theme for the plots in this article and can be
omitted.

```{r}
library(ekioplot)
color_palette <- ekioplot::ekio_pal()
theme_series <- function() {
  ekioplot::theme_ekio(base_size = 10) +
    theme(
      palette.color.discrete = color_palette
    )
}
```

## Construction costs and sentiment

The `fgv_ibre` dataset gathers indicators from Fundação Getulio Vargas (FGV).
FGV runs a wide range of surveys, and `realestatebr` collects the ones most
relevant to construction and real estate. The `name_series` and `code_series`
columns identify each series.

```{r}
fgv <- get_dataset("fgv_ibre")

count(fgv, name_series, name_simplified)
```

The plot below shows the confidence indicators of the construction sector. ICST
is the overall index, which averages the IE-CST (expectations) and the ISA-CST
(current situation). Although these series are already seasonally adjusted, we
smooth them with a 12-month moving average to make them more stable. As with
most confidence indicators, the values are normalized: values above 100
indicate a more favorable assessment, while values below 100 indicate a less
favorable assessment.

```{r}
confidence <- fgv |>
  filter(name_simplified %in% c("ic_cst", "ie_cst", "isa_cst")) |>
  augment_trends(group_cols = "name_simplified", methods = "ma") |>
  mutate(
    indicator = case_when(
      name_simplified == "ic_cst" ~ "ICST",
      name_simplified == "ie_cst" ~ "IE-CST (Expectations)",
      name_simplified == "isa_cst" ~ "ISA-CST (Current Situation)"
    )
  )

ggplot(confidence, aes(date, trend_ma)) +
  geom_line(aes(color = indicator), lwd = 0.8) +
  geom_hline(yintercept = 100) +
  labs(
    title = "Confidence Indicators",
    subtitle = "Construction sector confidence indicators, 12-month moving average.",
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

## Launches and sales

The `indicator` table from `abrainc` reports monthly launches, sales,
deliveries, cancellations, and supply for a panel of large developers. The
series are volatile, so the plot below uses values accumulated over twelve
months.

```{r}
abrainc <- get_dataset("abrainc", table = "indicator")

glimpse(abrainc)
```

```{r}
abrainc_trends <- abrainc |>
  filter(
    category %in% c("new_units", "sold"),
    variable == "total"
  ) |>
  mutate(
    units_12m = zoo::rollsumr(value, k = 12, fill = NA) / 1e3,
    category_label = case_when(
      category == "new_units" ~ "Launches",
      category == "sold" ~ "Sales",
      TRUE ~ NA_character_
    )
  ) |>
  trendseries::augment_trends(
    value_col = "units_12m",
    group_cols = "category_label"
  )

ggplot(abrainc_trends, aes(date, units_12m, color = category_label)) +
  geom_line(lwd = 0.5, alpha = 0.5) +
  geom_line(aes(y = trend_stl), lwd = 0.8) +
  facet_wrap(vars(category_label), ncol = 1) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "Launches and Sales in the Primary Market",
    subtitle = "Units accumulated in 12 months, ABRAINC-FIPE panel",
    y = "Units (thousand)",
    color = NULL
  ) +
  theme_series()
```

## Market segments

Brazil has two large segments: (1) social housing production, which is
financed and subsidized by the "Minha Casa Minha Vida" program (usually
abbreviated as "MCMV"); and (2) market-rate developments. The ABRAINC dataset
reports sales and launches for each of these markets.


```{r}
segments <- abrainc |>
  filter(
    category == "sold",
    variable %in% c("market_rate", "social_housing")
  ) |>
  summarise(
    year_total = if (anyNA(value)) NA_real_ else sum(value) / 1e3,
    .by = c("variable_label", "year")
  ) |>
  filter(!is.na(year_total))

ggplot(segments, aes(year, year_total)) +
  geom_col(aes(fill = variable_label)) +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 8)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    title = "Sales by Market Segment",
    subtitle = "Total yearly sales by market segment.",
    y = "Units (thousand)",
    fill = NULL
  ) +
  theme_series()
```

## Business conditions

The `radar` table condenses the business environment into standardized 0-10
scores across four dimensions: (1) macroeconomy; (2) credit; (3) demand; and
(4) the real estate sector. The plot shows one representative indicator from
each dimension, smoothed with a six-month moving average. Note that the moving
average is already bundled with the data in the `ma6` column.

```{r}
radar <- get_dataset("abrainc", table = "radar")

radar_sel <- radar |>
  filter(
    variable %in%
      c("confidence", "finance_condition", "employment", "input_costs"),
    date >= as.Date("2012-01-01")
  )

ggplot(radar_sel, aes(date, ma6)) +
  geom_line(color = color_palette[1], lwd = 0.8) +
  geom_hline(yintercept = 5, linetype = 2, color = "gray40") +
  facet_wrap(vars(variable_label)) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(
    title = "ABRAINC-FIPE Radar",
    subtitle = "Standardized scores (0-10), 6-month moving average.",
    y = "Score"
  ) +
  theme_series()
```

The radar indicator has been discontinued upstream, so the series ends at its
last published month. The [GitHub
issue](https://github.com/viniciusoike/realestatebr/issues/6) tracks its
status.

## São Paulo

São Paulo is Brazil's largest real estate market, and two sources cover it
here. The `leading` table from `abrainc` reports building permits in the city,
while `secovi` reports supply and sales for the city and the metropolitan
region.

### A leading indicator

Building permits anticipate construction activity. The `leading` table
compiles permits in the city of São Paulo into a leading indicator of real
estate activity.

```{r}
leading <- get_dataset("abrainc", table = "leading")

leading_index <- leading |>
  filter(
    variable == "leading_index",
    zone != "Total",
    date >= as.Date("2010-01-01")
  )

ggplot(leading_index, aes(date, value)) +
  geom_line(color = color_palette[1], lwd = 0.8) +
  facet_wrap(vars(zone)) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "Real Estate Leading Indicator",
    subtitle = "Based on building permits in São Paulo (100 = Dec/2000)",
    y = "Index"
  ) +
  theme_series()
```


### Supply and sales

The `launch` and `sale` tables from `secovi` track new supply and sales in the
city.

```{r}
secovi <- get_dataset("secovi")

count(secovi, category, variable)
```

```{r}
sp_market <- secovi |>
  filter(
    variable %in% c("launches", "sales"),
    name == "unidades"
  ) |>
  mutate(
    units_12m = zoo::rollsumr(value, k = 12, fill = NA) / 1e3
  ) |>
  trendseries::augment_trends(
    value_col = "units_12m",
    group_cols = "variable"
  ) |>
  mutate(
    variable_label = if_else(variable == "launches", "Launches", "Sales")
  )

ggplot(sp_market, aes(date)) +
  geom_line(
    aes(y = units_12m, color = variable_label),
    lwd = 0.6,
    alpha = 0.5
  ) +
  geom_line(aes(y = trend_stl, color = variable_label), lwd = 0.8) +
  facet_wrap(vars(variable_label), ncol = 1) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  guides(color = "none") +
  labs(
    title = "Launches and Sales in São Paulo",
    subtitle = "Units accumulated in 12 months, SECOVI-SP",
    y = "Units (thousand)",
    color = NULL
  ) +
  theme_series()
```

Some series break the totals down further. The plot below shows sales by
number of rooms.

```{r}
sp_market_rooms <- secovi |>
  filter(
    grepl("sales_[0-9]rooms", variable),
    name == "unidades"
  ) |>
  trendseries::augment_trends(group_cols = "variable") |>
  mutate(
    variable_label = regmatches(variable, regexpr("\\d+", variable)),
    variable_label = if_else(
      variable == "sales_1rooms",
      paste(variable_label, "room"),
      paste(variable_label, "rooms")
    )
  )

ggplot(sp_market_rooms, aes(date)) +
  geom_line(aes(y = trend_stl, color = variable_label), lwd = 0.8) +
  facet_wrap(vars(variable_label)) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  guides(color = "none") +
  labs(
    title = "Sales in São Paulo",
    subtitle = "STL trends for sales by number of rooms.",
    y = "Units (thousand)"
  ) +
  theme_series()
```

## Learn more

The help topics `?abrainc`, `?secovi`, and `?fgv_ibre` document the columns of
each dataset used here. For housing credit, see [Housing Credit in
Brazil](housing-credit.html); for price indices, see [Working with Property
Price Indices](working-with-rppi.html).
