---
title: "realestatebr 1.0.1"
description: >
  The first CRAN release of realestatebr brings a single, unified interface
  for Brazilian real estate data.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

I am happy to announce that `realestatebr` 1.0.1 is the first version of the
package on CRAN. The goal of `realestatebr` is to provide a unified interface to
Brazilian real estate data from many public sources, delivering everything in a
tidy `tibble` format.

```{r}
install.packages("realestatebr")
```

This release marks a milestone in how the package is used. The headline change
is the move from a long list of individual `get_*()` functions to a single entry
point, `get_dataset()`.

## One function instead of many

Earlier versions exposed a separate function for each source. To read financing
data from ABECIP you called `get_abecip_indicators()`, for the Central Bank
series you called `get_bcb_series()`, and so on. Every source had its own
function, its own arguments, and its own quirks.

```{r}
# The old way, no longer available
abecip <- get_abecip_indicators(table = "sbpe")
bcb <- get_bcb_series(table = "price")
secovi <- get_secovi()
```

Separate functions helped with discoverability and documentation, but they
stopped scaling as the number of datasets grew. From 1.0.1 onwards,
`get_dataset()` retrieves any dataset in the package.

```{r}
library(realestatebr)

# The new way
abecip <- get_dataset("abecip", table = "sbpe")
bcb <- get_dataset("bcb_series", table = "core")
secovi <- get_dataset("secovi", table = "all")
```

When the `table` argument is omitted, `get_dataset()` returns the default table
for that dataset.

```{r}
# Returns the default table
abecip <- get_dataset("abecip")
```

The individual `get_*()` functions are no longer part of the public interface.
The table below maps the most common ones to their `get_dataset()` equivalent.

| Old function | New call |
|----|----|
| `get_abecip_indicators("sbpe")` | `get_dataset("abecip", "sbpe")` |
| `get_abrainc_indicators()` | `get_dataset("abrainc")` |
| `get_bcb_realestate()` | `get_dataset("bcb_realestate")` |
| `get_bcb_series("price")` | `get_dataset("bcb_series", "core")` |
| `get_secovi()` | `get_dataset("secovi")` |
| `get_rppi("fipezap")` | `get_dataset("rppi", "fipezap")` |
| `get_rppi_bis()` | `get_dataset("rppi_bis")` |

## Discovering what is available

Since the names of datasets and tables are now arguments, you need a way to find
them. `list_datasets()` returns a `tibble` describing every dataset, its source,
and the tables it contains.

```{r}
datasets <- list_datasets()
datasets
```

The currently active datasets cover financing and credit, launches and sales,
macroeconomic series, and a wide range of residential property price indices.

| Dataset | Source | Tables |
|----|----|----|
| `abecip` | ABECIP | `sbpe`, `units`, `cgi` |
| `abrainc` | ABRAINC / FIPE | `indicator`, `radar`, `leading` |
| `bcb_realestate` | Banco Central do Brasil | `accounting`, `application`, `indices`, `sources`, `units` |
| `bcb_series` | Banco Central do Brasil | `core`, `primary`, `secondary`, `tertiary`, `full` |
| `fgv_ibre` | FGV IBRE | — |
| `rppi` | FipeZap, IVG-R, IGMI-R, IQA, IVAR, SECOVI-SP | `sale`, `rent`, and individual indices |
| `rppi_bis` | Bank for International Settlements | `selected`, `detailed_monthly`, `detailed_quarterly` |
| `secovi` | SECOVI-SP | `condo`, `rent`, `launch`, `sale` |

Users of previous versions might miss the tables from CBIC (Câmara Brasileira
da Indústria da Construção) and from Registro de Imóveis (`property_records`).
Registro de Imóveis no longer publishes consolidated Excel reports; the data
now comes in a different format, which the package does not yet read. CBIC
[paywalled](https://cbic.org.br/hubdedados/) all of its previously open
datasets. Several CBIC tables were built on open sources, so they may return
in a future release.

## Choosing where data comes from

`get_dataset()` resolves data in two tiers. By default it reads a pre-processed
asset from the package's GitHub release, which is fast and updated automatically
by a weekly pipeline. If that asset is unavailable, it falls back to a fresh
download from the original source. The `source` argument lets you control this
behaviour.

```{r}
# Auto (default): GitHub release, then fresh download as a fallback
data <- get_dataset("abecip")

# Pre-processed asset from the package's GitHub release
data <- get_dataset("abecip", source = "github")

# Fresh download straight from the original source
data <- get_dataset("abecip", source = "fresh")
```

Repeated calls within one R session are served from an in-memory store, so
asking for the same dataset twice does not download it twice. Use
`clear_session_cache()` to drop that store without restarting R.

```{r}
clear_session_cache()
```

This release also reworks the caching architecture so that the package never
writes outside the R session's temporary directory, in line with CRAN policy.
The earlier user-level disk cache and its helper functions have been removed.

## A worked example

The example below reads the FipeZap index and plots the year-on-year change in
sale and rent prices for the city of São Paulo.

```{r}
library(dplyr)
library(ggplot2)

plot_font <- if (
  requireNamespace("systemfonts", quietly = TRUE) &&
    any(systemfonts::system_fonts()[["family"]] == "Avenir")
) {
  "Avenir"
} else {
  "sans"
}

fipezap <- get_dataset("rppi", table = "fipezap")

rppi_spo <- fipezap |>
  filter(
    name_muni == "São Paulo",
    market == "residential",
    rooms == "total",
    variable == "acum12m",
    date >= as.Date("2019-01-01")
  )

ggplot(rppi_spo, aes(x = date, y = value, color = rent_sale)) +
  geom_line(lwd = 0.8) +
  geom_hline(yintercept = 0) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "São Paulo Property Price Index",
    x = NULL,
    y = "YoY chg. (%)",
    color = ""
  ) +
  theme_minimal(base_family = plot_font) +
  theme(legend.position = "bottom")
```

## Breaking changes

Consolidating the interface meant retiring a few datasets that could not be
maintained reliably. The `cbic`, `property_records`, `nre_ire`, and `itbi`
datasets have been removed. `property_records` and `cbic` may return, as noted
above, but `nre_ire` and `itbi` will not. `nre_ire` resists automation and
costs more upkeep than it returns. `itbi` was an experiment in bringing large
transaction-level datasets into the package; that scale of data belongs in a
package of its own.

The `bcb_series` and `rppi` datasets also changed shape. `bcb_series` now
returns a compact set of columns and accepts a hierarchy level such as `"core"`
or `"primary"` instead of a category name, while the stacked `rppi` table gains
`transaction_type` and `source` columns. See the
[changelog](../news/index.html) for the full list.

## Learn more

To go further, read the
[Getting Started](getting-started.html) article for a tour of the core
interface, or [Working with RPPI](working-with-rppi.html) for a deeper look at
the property price indices. Bug reports and suggestions are welcome on
[GitHub](https://github.com/viniciusoike/realestatebr/issues).
