---
title: "Working with MCMV"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  message = FALSE,
  warning = FALSE
)
```

The Ministério das Cidades publishes open data on Minha Casa, Minha Vida
(MCMV) as three files. `realestatebr` normalizes them into three lazy tables
and queries them with DuckDB, so filters and aggregations run before any data
enter R memory. The financing table alone holds almost eight million rows.

The data are published by the Secretaria Nacional de Habitação (SNH) under a
[Creative Commons Attribution](https://dadosabertos.cidades.gov.br/dataset/dados-do-programa-minha-casa-minha-vida-pmcmv)
license. Cite the Ministério das Cidades when you use them.

| Table | Grain |
|---|---|
| `financing` | One published financing contract, usually one housing unit |
| `financing_summary` | Official totals by municipality, year, month, and income band |
| `subsidized_projects` | One published project record financed by the federal budget (OGU) |

The [MCMV data dictionary](mcmv-data-dictionary.html) defines every column
and documents how the source files changed across releases.

## Open the dataset

Install the optional query dependencies once.

```{r install}
install.packages(c("DBI", "dbplyr", "duckdb"))
```

Open all three tables through one catalog.

```{r open}
library(realestatebr)
library(dplyr)

mcmv <- query_dataset("mcmv")
mcmv
```

## Aggregate before collecting

Each `financing` row reports one financed unit. Summing `units_financed`
therefore counts contracts, and summing `amount_financed` gives the nominal
credit volume.

```{r yearly}
yearly <- mcmv$financing |>
  mutate(year = year(contract_date)) |>
  group_by(year) |>
  summarise(
    units = sum(units_financed, na.rm = TRUE),
    amount_bn = sum(amount_financed, na.rm = TRUE) / 1e9
  ) |>
  arrange(year) |>
  collect()

tail(yearly, 4)
#> # A tibble: 4 × 3
#>    year  units amount_bn
#>   <dbl>  <dbl>     <dbl>
#> 1  2023 491209      74.4
#> 2  2024 605554      99.2
#> 3  2025 668553     116.
#> 4  2026 398641      74.5
```

The last year is partial: it ends at the snapshot's reference date. Amounts
are nominal BRL. Deflate them before comparing years.

Filter on `state` or `code_muni_6` to restrict the query to a region. The
financing Parquet file is sorted by state and municipality, so DuckDB reads
only the relevant parts of it.

```{r municipalities}
mcmv$financing |>
  filter(state == "SP", contract_date >= as.Date("2023-01-01")) |>
  group_by(name_muni) |>
  summarise(
    units = sum(units_financed, na.rm = TRUE),
    amount_mn = sum(amount_financed, na.rm = TRUE) / 1e6
  ) |>
  arrange(desc(units)) |>
  head(3) |>
  collect()
#> # A tibble: 3 × 3
#>   name_muni       units amount_mn
#>   <chr>           <dbl>     <dbl>
#> 1 SÃO PAULO      250441    50224.
#> 2 SOROCABA        18421     3211.
#> 3 RIBEIRÃO PRETO  18378     3153.
```

## Income bands are source codes

Recent releases publish income bands as codes `1` to `4`, without labels, and
the official dictionary does not define them. `realestatebr` keeps the codes
as published instead of guessing a crosswalk. Code `4` appears only in
contracts signed from 2025 onward.

```{r income-band}
mcmv$financing_summary |>
  filter(contract_year >= 2025) |>
  group_by(contract_year, income_band) |>
  summarise(units = sum(units_financed, na.rm = TRUE), .groups = "drop") |>
  arrange(contract_year, income_band) |>
  collect()
#> # A tibble: 10 × 3
#>    contract_year income_band  units
#>            <int> <chr>        <dbl>
#>  1          2025 1           238421
#>  2          2025 2           194023
#>  3          2025 3           199630
#>  4          2025 4            29969
#>  5          2025 <NA>          6510
#>  6          2026 1           132237
#>  7          2026 2            92587
#>  8          2026 3           147466
#>  9          2026 4            25341
#> 10          2026 <NA>          1010
```

A missing income band is not the same as a contract outside MCMV. The
`financing_program` column carries the source's program labels, including
`Fundo Social` and `Classe Média`.

## Use the summary for aggregate series

`financing_summary` holds the Ministry's own totals. For the July 2026
release, these totals agree with the sums of `financing` in every
municipality, month, and income band, to within BRL 0.01. The summary is
about ten times smaller, so it answers aggregate questions faster.

`subsidy_total` in the summary equals the sum of all four subsidy columns in
`financing`: the FGTS and OGU discounts plus the FGTS and OGU interest
subsidies. This relationship was checked for the July 2026 release; verify it
again when using a different snapshot.

## Harmonize labels in your query

The package preserves source labels, including inconsistent capitalization.
`property_type`, for instance, contains `novo`, `Novo`, and `NOVO`.
Harmonize labels inside the query when your analysis needs it.

```{r labels}
mcmv$financing |>
  mutate(property_type = lower(property_type)) |>
  count(property_type) |>
  collect()
#> # A tibble: 3 × 2
#>   property_type       n
#>   <chr>           <dbl>
#> 1 usado         1504065
#> 2 <NA>             6565
#> 3 novo          6339252
```

`fgts_account_holder` mixes `S`/`N` with `1`/`0` and is missing in most
rows. It is stored as text, not as a logical column.

## Subsidized projects contain duplicates

`subsidized_projects` includes exact duplicate rows, and `operation_code` is
neither unique nor always present. Remove exact duplicates before counting
projects or units.

```{r projects}
mcmv$subsidized_projects |>
  distinct() |>
  group_by(project_status) |>
  summarise(
    projects = n(),
    units_contracted = sum(units_contracted, na.rm = TRUE),
    units_delivered = sum(units_delivered, na.rm = TRUE),
    .groups = "drop"
  ) |>
  collect()
#> # A tibble: 3 × 4
#>   project_status       projects units_contracted units_delivered
#>   <chr>                   <dbl>            <dbl>           <dbl>
#> 1 Concluído               15588          1648140         1646858
#> 2 Distratado/Cancelado      395            48268             686
#> 3 Não Concluído            6847           410701           36606
```

`distinct()` removes only identical rows. Records that repeat an operation
code with different values remain, and the source gives no rule for choosing
between them.

The unit columns satisfy `units_contracted = units_delivered +
units_outstanding + units_cancelled` in every row where all four are
reported. A few rows lack `units_cancelled`; the package keeps it missing
rather than filling in zero.

## Joining tables

The tables share no contract or project identifier. Do not join `financing`
to `subsidized_projects` by project name. Compare them only as aggregates,
for instance by municipality and year, and keep in mind that they cover
different programs: `financing` covers FGTS-funded credit, while
`subsidized_projects` covers projects funded by the federal budget.

`code_muni_6` is the six-digit IBGE municipality code, without the check
digit. To join with seven-digit codes, such as those in `geobr`, drop the last
digit of the seven-digit code.

## Snapshot versions

Every MCMV release replaces the full history, so the package publishes each
release as an immutable snapshot. `query_dataset("mcmv")` opens the latest
one. Pin a version when an analysis must be reproducible.

```{r version}
mcmv_2026_07 <- query_dataset("mcmv", version = "2026-07-24")
close(mcmv_2026_07)
```

Close the catalog after the final query. Lazy tables obtained from it become
invalid after `close()`.

```{r close}
close(mcmv)
```
