---
title: "Working with PNAD Housing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with PNAD Housing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`pnad_housing` provides annual household estimates from five PNAD Contínua
SIDRA tables. The default table describes housing tenure.

```{r, eval=FALSE}
tenure <- get_dataset("pnad_housing")

rented_share <- tenure |>
  dplyr::filter(
    geography_type == "brazil",
    category == "rented",
    variable == "household_share"
  )
```

Counts remain in thousand households. Shares and coefficients of variation
remain percentages on a 0-100 scale. Each table includes its total category;
exclude `category == "total"` before summing component categories.

```{r, eval=FALSE}
dwelling_type <- get_dataset("pnad_housing", table = "dwelling_type")

apartment_share <- dwelling_type |>
  dplyr::filter(
    geography_type == "brazil",
    category == "apartment",
    variable == "household_share"
  )
```

The housing-module tables publish 2016-2019 and 2022 onward. They contain no
2020 or 2021 observations. Domestic-unit composition has a longer annual
series and describes relationships to the reference person; it is not a count
of families.

```{r, eval=FALSE}
household_size <- get_dataset("pnad_housing", table = "household_size")

one_person <- household_size |>
  dplyr::filter(
    geography_type == "brazil",
    category == "one_resident",
    variable %in% c("households", "household_share")
  )

mean_size <- get_dataset("pnad_housing", table = "mean_household_size") |>
  dplyr::filter(
    geography_type == "brazil",
    variable == "mean_residents"
  )
```

The composition table publishes counts and their CVs. A composition share can
be calculated from a category count and the same-table total, but the count CV
is not the CV of that derived share.
