---
title: "Working with CNO"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  message = FALSE,
  warning = FALSE
)
```

The Cadastro Nacional de Obras (CNO) is a relational dataset. One table
describes each registered construction work, while three child tables record
areas, CNAE activities, and responsibility periods. `realestatebr` queries
these tables lazily with DuckDB, so filtering and aggregation happen before
data enter R memory.

## Open the dataset

Install the optional query dependencies once.

```{r install}
install.packages(c("DBI", "dbplyr", "duckdb"))
```

Open all related tables through one catalog.

```{r open}
library(realestatebr)
library(dplyr)

cno <- query_dataset("cno")

cno
```

The catalog contains four lazy tables.

| Table | Grain |
|---|---|
| `constructions` | One row per CNO registration |
| `areas` | One row per reported area classification |
| `cnaes` | One row per CNAE registration |
| `responsibilities` | One row per responsibility period |

All four tables use the twelve-digit character column `cno` as their join
key. Keeping the tables in the same catalog ensures that joins run inside one
DuckDB connection.

## Filter before collecting

The object returned by a dplyr pipeline remains lazy. `collect()` executes the
query and returns its result as an in-memory tibble.

```{r filter}
constructions_sp <- cno$constructions |>
  filter(
    state == "SP",
    registration_date >= as.Date("2025-01-01")
  ) |>
  select(cno, municipality_name, registration_date, status_code)

constructions_sp

constructions_sp |>
  collect()
```

Select only the columns needed by the analysis. DuckDB can then avoid reading
unused Parquet columns over the network.

## Join constructions and areas

Each construction can have several area rows. The following query returns residential
area records for construction works in São Paulo.

```{r join-areas}
residential_areas <- cno$areas |>
  filter(destination == "Residencial unifamiliar") |>
  select(cno, construction_category, structure_type, area_type, area)

sp_residential <- constructions_sp |>
  inner_join(residential_areas, by = "cno") |>
  collect()
```

The source may repeat a work's total area across several classifications, and
it contains extreme area values. The package preserves these source records.
Do not assume that summing every `areas$area` row produces an unbiased measure
of built area.

Let's look at CNO `"900143355070"`. This case shows how exact duplicate rows
affect a join between the `constructions` and `areas` tables.

```{r join-areas-duplicates}
construction_case <- cno$constructions |>
  filter(cno == "900143355070") |>
  select(cno, municipality_name, state, total_area)

area_case <- cno$areas |>
  filter(cno == "900143355070")

case <- construction_case |>
  inner_join(area_case, by = "cno") |>
  collect()

case |>
  select(
    cno,
    municipality_name,
    total_area,
    destination,
    area_type,
    complementary_area_type,
    area
  )

# A tibble: 8 × 7
#   cno          municipality_name total_area destination               area_type    complementary_area_type            area
#   <chr>        <chr>                  <dbl> <chr>                     <chr>        <chr>                             <dbl>
# 1 900143355070 LONDRINA              18719. Residencial multifamiliar Complementar Quadra Esportiva e Poliesportiva   160
# 2 900143355070 LONDRINA              18719. Residencial multifamiliar Complementar Estacionamento Térreo             3076.
# 3 900143355070 LONDRINA              18719. Residencial multifamiliar Complementar Piscina                            136.
# 4 900143355070 LONDRINA              18719. Residencial multifamiliar Principal    NA                               15347.
# 5 900143355070 LONDRINA              18719. Residencial multifamiliar Complementar Quadra Esportiva e Poliesportiva   160
# 6 900143355070 LONDRINA              18719. Residencial multifamiliar Complementar Estacionamento Térreo             3076.
# 7 900143355070 LONDRINA              18719. Residencial multifamiliar Complementar Piscina                            136.
# 8 900143355070 LONDRINA              18719. Residencial multifamiliar Principal    NA                               15347.
```

These duplicates are present in the original CNO `areas` dataset. `realestatebr`
currently reproduces them exactly. For most use cases, these duplicate rows
will lead to errors, so we advise removing them using, for instance, `dplyr::distinct()`.

Removing the duplicate rows, in this case, would also make the sum of the individual
`area` column from `areas` coincide with `total_area` from `constructions`. This
will not always be the case.

```{r remove-area-duplicates}
construction_case <- cno$constructions |>
  filter(cno == "900143355070") |>
  select(cno, municipality_name, state, total_area)

area_case <- cno$areas |>
  filter(cno == "900143355070") |>
  distinct()

case <- construction_case |>
  inner_join(area_case, by = "cno") |>
  collect()

all.equal(unique(case$total_area), sum(case$area))
#> [1] TRUE
```

Future versions of `realestatebr` may add validation steps that remove exact
duplicate rows and other noise from these tables.

## Combine several child tables safely

Joining `areas` and `cnaes` directly can multiply rows. A work with two area
rows and three CNAE rows produces six joined rows. Aggregate each child table
to one row per CNO before combining them when the analysis needs one row per
work.

```{r aggregate-before-join}
area_summary <- cno$areas |>
  group_by(cno) |>
  summarise(
    area_records = n(),
    reported_area = sum(area, na.rm = TRUE),
    .groups = "drop"
  )

cnae_summary <- cno$cnaes |>
  group_by(cno) |>
  summarise(
    cnae_records = n(),
    .groups = "drop"
  )

construction_summary <- cno$constructions |>
  filter(state == "SP") |>
  left_join(area_summary, by = "cno") |>
  left_join(cnae_summary, by = "cno") |>
  collect()
```

The appropriate area aggregation depends on the research question. `sum()`
above demonstrates the relational pattern; it is not a general cleaning rule.

## Municipality identifiers

`municipality_tom_code` is Receita Federal's four-digit TOM code. It is not
the seven-digit IBGE municipality code used by `geobr` and most Brazilian
regional datasets. A TOM-to-IBGE crosswalk is required before joining those
sources.

## Raw values and snapshot versions

The tables normalize headers, dates, numeric fields, identifiers, and text
encoding. They otherwise preserve the source. This includes unusual state
values, historical date sentinels, area outliers, status codes, and reported
category labels.

`query_dataset("cno")` resolves `"latest"` once and pins that immutable
snapshot while the catalog remains open. Supply a published version explicitly
when an analysis must be reproducible.

```{r version}
cno_2026_09_13 <- query_dataset("cno", version = "2026-09-13")
close(cno_2026_09_13)
```

Close the catalog after the final query. Any lazy tables obtained from it
become invalid after `close()`.

```{r close}
close(cno)
```
