---
title: "Working with CNO"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  message = FALSE,
  warning = FALSE
)
```

The Cadastro Nacional de Obras (CNO) is a relational dataset. One table
describes each registered construction work, while three child tables record
areas, CNAE activities, and responsibility periods. `realestatebr` queries
these tables lazily with DuckDB, so filtering and aggregation happen before
data enter R memory.

## Open the dataset

Install the optional query dependencies once.

```{r install}
install.packages(c("DBI", "dbplyr", "duckdb"))
```

Open all related tables through one catalog.

```{r open}
library(realestatebr)
library(dplyr)

cno <- query_dataset("cno")

cno
```

The catalog contains four lazy tables.

| Table | Grain |
|---|---|
| `works` | One row per CNO registration |
| `areas` | One row per reported area classification |
| `cnaes` | One row per CNAE registration |
| `responsibilities` | One row per responsibility period |

All four tables use the twelve-digit character column `cno` as their join
key. Keeping the tables in the same catalog ensures that joins run inside one
DuckDB connection.

## Filter before collecting

The object returned by a dplyr pipeline remains lazy. `collect()` executes the
query and returns its result as an in-memory tibble.

```{r filter}
works_sp <- cno$works |>
  filter(
    state == "SP",
    registration_date >= as.Date("2025-01-01")
  ) |>
  select(cno, municipality_name, registration_date, status_code)

works_sp

works_sp |>
  collect()
```

Select only the columns needed by the analysis. DuckDB can then avoid reading
unused Parquet columns over the network.

## Join works and areas

Each work can have several area rows. The following query returns residential
area records for works in São Paulo.

```{r join-areas}
residential_areas <- cno$areas |>
  filter(destination == "Residencial unifamiliar") |>
  select(cno, work_category, structure_type, area_type, area)

sp_residential <- works_sp |>
  inner_join(residential_areas, by = "cno") |>
  collect()
```

The source may repeat a work's total area across several classifications, and
it contains extreme area values. The package preserves these source records.
Do not assume that summing every `areas$area` row produces an unbiased measure
of built area.

## Combine several child tables safely

Joining `areas` and `cnaes` directly can multiply rows. A work with two area
rows and three CNAE rows produces six joined rows. Aggregate each child table
to one row per CNO before combining them when the analysis needs one row per
work.

```{r aggregate-before-join}
area_summary <- cno$areas |>
  group_by(cno) |>
  summarise(
    area_records = n(),
    reported_area = max(area, na.rm = TRUE),
    .groups = "drop"
  )

cnae_summary <- cno$cnaes |>
  group_by(cno) |>
  summarise(
    cnae_records = n(),
    .groups = "drop"
  )

work_summary <- cno$works |>
  filter(state == "SP") |>
  left_join(area_summary, by = "cno") |>
  left_join(cnae_summary, by = "cno") |>
  collect()
```

The appropriate area aggregation depends on the research question. `max()`
above demonstrates the relational pattern; it is not a general cleaning rule.

## Municipality identifiers

`municipality_tom_code` is Receita Federal's four-digit TOM code. It is not
the seven-digit IBGE municipality code used by `geobr` and most Brazilian
regional datasets. A TOM-to-IBGE crosswalk is required before joining those
sources.

## Raw values and snapshot versions

The tables normalize headers, dates, numeric fields, identifiers, and text
encoding. They otherwise preserve the source. This includes unusual state
values, historical date sentinels, area outliers, status codes, and reported
category labels.

`query_dataset("cno")` resolves `"latest"` once and pins that immutable
snapshot while the catalog remains open. Supply a published version explicitly
when an analysis must be reproducible.

```{r version}
cno_2026_09_13 <- query_dataset("cno", version = "2026-09-13")
close(cno_2026_09_13)
```

Close the catalog after the final query. Any lazy tables obtained from it
become invalid after `close()`.

```{r close}
close(cno)
```
