---
title: "Working with Property Price Indices"
---

```{r knitr-setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618,
  fig.align = "center",
  message = FALSE,
  warning = FALSE
)
```

## Introduction

Brazil has several residential property price indices (RPPIs), each built by a
different institution and each with its own methodology, geographic coverage,
and time span. This article shows how to reach every one of them through
`get_dataset("rppi")`. Since `realestatebr` returns tibbles, we recommend using
it together with `dplyr`.

```{r load-packages}
library(realestatebr)
library(dplyr)
```

The code below defines an optional common theme for the plots in this article.
It can be omitted.

```{r setup, message = FALSE}
library(ggplot2)

color_palette <- c(
  "#1E3A5F",
  "#DD6B20",
  "#2C7A7B",
  "#D69E2E",
  "#805AD5",
  "#C53030"
)

theme_series <- function() {
  theme_minimal(base_size = 10) +
    theme(
      plot.title = element_text(size = 16),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.line.x = element_line(color = "gray10", linewidth = 0.5),
      axis.ticks.x = element_line(color = "gray10", linewidth = 0.5),
      axis.title.x = element_blank(),
      legend.position = "bottom",
      palette.color.discrete = color_palette
    )
}
```

## Choosing an index

The IGMI-R is a good default for sale prices and the IVAR for rent prices.
FipeZap is the main option for commercial prices. Because the indices draw on
different source data and methods, the right choice depends on the question.

The IGMI-R is a hedonic price index based on transaction prices from housing
mortgages. Its limits are the short time span (2014-present) and the coverage
of major cities only.

The IVAR is a repeat-rent index based on residential rental contracts. It also
starts late (2019) and covers four cities.

FipeZap has the widest geographic (50+ cities) and temporal coverage
(2008-present), but rests on median listing prices rather than transaction
prices. It answers questions about advertised prices, not about what buyers
actually paid.

### Working with realestatebr

To get a dataset, use `get_dataset()`

```{r load-datasets-overview}
#| eval: false
igmi <- get_dataset("rppi", table = "igmi")
ivar <- get_dataset("rppi", table = "ivar")
fipezap <- get_dataset("rppi", table = "fipezap")
```

For easier comparison between indices, the `table` parameter accepts `"sale"`
and `"rent"` as options, which stack the available indices into a single
`tibble`.

```{r load-stacked-datasets}
#| eval: false
sale <- get_dataset("rppi", table = "sale")
rent <- get_dataset("rppi", table = "rent")
```

## Sale indices

### IVG-R

The IVG-R (Índice de Valores de Garantia de Imóveis Residenciais) is published
by the Brazilian Central Bank. It covers the national market from 2001 and is
widely regarded as the official RPPI in Brazil. The index is based on median
appraisal prices from mortgage contracts, smoothed with a 3-month moving
average and the HP filter.

```{r ivgr-load}
ivgr <- get_dataset("rppi", "ivgr")

glimpse(ivgr)
```

```{r ivgr-plot}
ggplot(ivgr, aes(date, index)) +
  geom_line(color = color_palette[1], linewidth = 0.7) +
  labs(
    title = "IVG-R — National Sale Index",
    x = NULL,
    y = "Index"
  ) +
  theme_series()
```

### IGMI-R

The IGMI-R (Índice Geral do Mercado Imobiliário Residencial) is published by
Abecip/FGV. It fits a hedonic pricing model to transaction prices, so it
tracks closed deals rather than asking prices. Coverage spans major Brazilian
cities from 2014.

```{r igmi-load}
igmi <- get_dataset("rppi", "igmi")

glimpse(igmi)
```

```{r igmi-plot}
main_cities <- c("São Paulo", "Rio De Janeiro", "Belo Horizonte", "Brasília")

subigmi <- igmi |>
  filter(name_muni %in% main_cities)

ggplot(subigmi, aes(date, index, color = name_muni)) +
  geom_line(linewidth = 0.8) +
  labs(
    title = "IGMI-R — Sale Index by City",
    x = NULL,
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

### FipeZap (sale)

FipeZap covers more than 50 cities, more than any other Brazilian RPPI, and is
based on median **listing prices** rather than transaction prices. It runs from
2008 for both residential and commercial markets, with breakdowns by number of
bedrooms. Prices are stratified by region and income level using Census
weights, and the final series is smoothed with a 3-month moving average.

Four columns are specific to FipeZap.

- `market` (residential/commercial)
- `rent_sale` (rent/sale)
- `rooms` (1, 2, 3, 4+, total)
- `variable` (index, chg, acum12m, price_m2, yield)

```{r fipezap-sale-load}
fz <- get_dataset("rppi", table = "fipezap")

glimpse(fz)
```

Working with this dataset requires more filtering than the others. For most
uses the three choices that matter are `market` ('residential' or
'commercial'), `rent_sale` ('sale' or 'rent'), and the desired `variable`.

```{r fipezap-sale-filter}
subzap <- fz |>
  filter(
    market == "residential",
    rent_sale == "sale",
    rooms == "total",
    variable == "index",
    name_muni %in% main_cities
  )
```

```{r fipezap-sale-plot}
ggplot(subzap, aes(date, value, color = name_muni)) +
  geom_line(linewidth = 0.8) +
  labs(
    title = "FipeZap — Residential Sale Index",
    x = NULL,
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

### Comparing sale indices

The `table` argument also accepts `"sale"`, which stacks the available sale
indices into a single `tibble`.

```{r sale-stacked-load}
sale_indices <- get_dataset("rppi", "sale")

glimpse(sale_indices)
```

The stacked table keeps only `rooms == "total"` for FipeZap, and the
nationwide index appears under `name_muni == "Brazil"` rather than a city
name.

The plot below shows the three national sale indices. They tracked each other
closely until the pandemic, after which the gap between them widens.

```{r sale-stacked-filter}
comp_index <- sale_indices |>
  filter(name_muni == "Brazil", date >= as.Date("2015-01-01"))
```

```{r sale-stacked-plot}
ggplot(comp_index, aes(date, acum12m * 100, color = source)) +
  geom_line(linewidth = 0.7) +
  labs(
    title = "Comparing Sale Indices in Brazil",
    subtitle = "12-month accumulated change (%)",
    y = "%",
    color = NULL
  ) +
  theme_series()
```

## Rent indices

### IVAR

The IVAR (Índice de Variação de Aluguéis Residenciais) is published by FGV and
uses a repeat-rent methodology applied to actual rental contracts. It covers
major Brazilian cities from 2019 and is available at both city and national
level. Notably, the index has no smoothing, making it rather volatile.

```{r ivar-load}
ivar <- get_dataset("rppi", table = "ivar")

glimpse(ivar)
```

The series is noisy, so the plot adds a 5-month moving average from the
`trendseries` package.

```{r ivar-trend}
library(trendseries)

ivar_trend <- ivar |>
  filter(name_muni != "Brazil") |>
  augment_trends(
    value_col = "index",
    group_cols = "name_muni",
    methods = "ma",
    window = 5
  )
```

```{r ivar-plot}
ggplot(ivar_trend, aes(date, color = name_muni)) +
  geom_line(aes(y = index), lwd = 0.5, alpha = 0.5) +
  geom_line(aes(y = trend_ma), lwd = 0.7) +
  geom_hline(yintercept = 100) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "IVAR — City Rent Indices",
    subtitle = "Smoothed moving average (5-month window)",
    x = NULL,
    y = "Index"
  ) +
  theme_series()
```

### IQA and IQAIW

QuintoAndar published two successive rental indices for São Paulo and Rio de
Janeiro. The **IQA** (2019–mid-2023) was based on median contract prices. In
mid-2023 it rebranded to **IQAIW** (Índice QuintoAndar ImovelWeb) and adopted
a hedonic model incorporating both listing and contract prices. The two series
are not directly comparable due to the methodology break.

The original IQA compared median contract prices, stratified by region and
income level with Census weights, much like FipeZap. The IQAIW uses a hedonic
model (spatial GAM), double imputation, and a similar Census-based
stratification.

The IQA source data reports rent prices (R$/m²) rather than an index, so
building an index requires a base-100 normalization. The package ships an
`index` column already normalized to the first available observation.

```{r iqa-iqaiw-load}
iqa <- get_dataset("rppi", "iqa")
iqaiw <- get_dataset("rppi", "iqaiw")

glimpse(iqaiw)
```

```{r iqa-plot}
ggplot(iqa, aes(date, index, color = name_muni)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = 100) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "IQA — Rent Index",
    subtitle = "Index (2019/06 = 100)",
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

The newer IQAIW index covers more cities and, like FipeZap, breaks results
down by number of bedrooms (`rooms`).

```{r iqaiw-plot}
ggplot(
  subset(iqaiw, rooms == "total" & !is.na(acum12m)),
  aes(date, acum12m * 100, color = name_muni)
) +
  geom_line(linewidth = 0.7) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "IQAIW — Rent Index",
    subtitle = "Accumulated 12-month change (%)",
    y = "%",
    color = NULL
  ) +
  theme_series()
```

The package keeps the discontinued IQA for its history. Plotting both series
together shows where the methodology break falls.

```{r quintoandar-combine}
quintoandar <- bind_rows(
  list(
    "IQA" = mutate(iqa, rooms = "total"),
    "IQAIW" = iqaiw
  ),
  .id = "source"
)

quintoandar_spo <- quintoandar |>
  filter(name_muni == "São Paulo", rooms == "total")
```

```{r quintoandar-compare}
ggplot(quintoandar_spo, aes(date, index, color = source)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = 100) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "QuintoAndar Rent Index — São Paulo",
    subtitle = "IQA (pre-2023) and IQAIW (post-2023) use different methodologies",
    x = NULL,
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

### FipeZap (rent)

FipeZap provides the broadest geographic coverage for rental indices, but is
based on median listing prices rather than contracts. The example below shows
the 12-month accumulated change across cities with complete data from 2021.

```{r fipezap-rent-load}
fz <- get_dataset("rppi", table = "fipezap")

glimpse(fz)
```

This is the same table used in the sale example above; only the `rent_sale`
filter changes.

```{r fipezap-rent-filter}
fz_rent <- fz |>
  filter(
    market == "residential",
    rent_sale == "rent",
    rooms == "total",
    variable == "acum12m",
    date >= as.Date("2021-01-01")
  )

sel_cities <- fz_rent |>
  filter(date == as.Date("2021-01-01"), !is.na(value)) |>
  pull(name_muni)
```

The plot below uses every city with observations starting in 2021, which shows
how far FipeZap's geographic coverage reaches.

```{r fipezap-rent-plot}
#| fig-width: 12
#| fig-height: 9
#| out-width: 100%
ggplot(subset(fz_rent, name_muni %in% sel_cities), aes(date, value * 100)) +
  geom_line(linewidth = 0.7, color = color_palette[1]) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(name_muni)) +
  labs(
    title = "FipeZap — 12-month Rent Change by City",
    x = NULL,
    y = "Accumulated 12-month change (%)"
  ) +
  theme_series()
```

### SECOVI-SP

SECOVI-SP publishes a residential rent index for the city of São Paulo. The
series tracks median contract prices for residential leases and runs from
2008, making it one of the longest rent series for a single Brazilian city.

```{r secovi-load}
secovi <- get_dataset("rppi", "secovi_sp")

glimpse(secovi)
```

```{r secovi-plot}
ggplot(secovi, aes(date, acum12m * 100)) +
  geom_line(color = color_palette[1], linewidth = 0.7) +
  geom_hline(yintercept = 0) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "SECOVI-SP — Residential Rent Index",
    subtitle = "12-month accumulated change (%)",
    x = NULL,
    y = "%"
  ) +
  theme_series()
```

## Comparing indices

### Stacked tables

The `"sale"` and `"rent"` tables stack all available indices into a single
data frame, making cross-source comparisons straightforward.

```{r rent-stacked-load}
rent_indices <- get_dataset("rppi", "rent")
```

```{r rent-stacked-filter}
rent_indices_comp <- rent_indices |>
  filter(
    name_muni %in% c("São Paulo", "Rio de Janeiro"),
    date >= as.Date("2019-01-01")
  )
```

```{r rent-stacked-plot}
ggplot(rent_indices_comp, aes(date, acum12m * 100, color = source)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(name_muni)) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "Rent Indices — São Paulo and Rio de Janeiro",
    subtitle = "12-month accumulated change (%)",
    y = "Accumulated change (%)",
    color = NULL
  ) +
  theme_series()
```

### Normalizing to a common base

Each index has its own base period, so levels are not comparable until they
share one. The example below rebases the national sale indices to January
2018 = 100.

```{r sale-rebased-load}
sales <- get_dataset("rppi", "sale")

national <- sales |>
  filter(
    name_muni == "Brazil",
    date >= as.Date("2018-01-01"),
    date <= as.Date("2023-12-01")
  )

national_rebased <- national |>
  arrange(source, date) |>
  mutate(
    index_rebased = index / first(index) * 100,
    .by = source
  )

total_growth <- national_rebased |>
  summarise(
    growth = last(index_rebased) - first(index_rebased),
    date = last(date),
    index_rebased = last(index_rebased),
    .by = source
  ) |>
  mutate(label = sprintf("%s:\n+%.1f%%", source, growth))
```

```{r sale-rebased-plot}
ggplot(national_rebased, aes(date, index_rebased, color = source)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 100) +
  geom_label(
    data = total_growth,
    aes(label = label),
    hjust = 0,
    nudge_x = 30,
    show.legend = FALSE,
    size = 3
  ) +
  scale_x_date(
    date_breaks = "1 year",
    date_labels = "%Y",
    expand = expansion(mult = c(0, 0.125))
  ) +
  labs(
    title = "Brazil National Sale Indices — Rebased to Jan 2018",
    x = NULL,
    y = "Index (Jan 2018 = 100)",
    color = NULL
  ) +
  theme_series()
```

## International: BIS

The BIS dataset carries quarterly residential property price indices for 60+
countries, so you can compare Brazil against other markets.

```{r bis-load}
bis <- get_dataset("rppi_bis")

bis_sub <- bis |>
  filter(
    ref_area_name %in% c("Brazil", "United States", "Germany", "Japan"),
    is_nominal == 0,
    unit == "index",
    date >= as.Date("2000-01-01")
  )
```

```{r bis-plot}
ggplot(bis_sub, aes(date, value, color = ref_area_name)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 100, linetype = "dashed", alpha = 0.4) +
  labs(
    title = "Real Residential Property Prices",
    subtitle = "BIS, index 2010 = 100",
    x = NULL,
    y = "Index",
    color = NULL
  ) +
  theme_series()
```

## Index reference

| Index | Source | Coverage | Methodology | From |
|-------|--------|----------|-------------|------|
| IVG-R | BCB | National | Median appraisal prices, HP-filtered | 2001 |
| IGMI-R | ABECIP/FGV | Major cities | Hedonic (transaction prices) | 2014 |
| FipeZap | FIPE/ZAP | 50 cities | Median listing prices | 2008 |
| IVAR | FGV | Major cities + national | Repeat-rent (contracts) | 2019 |
| IQA | QuintoAndar | São Paulo, Rio | Median contract prices | 2019 |
| IQAIW | QuintoAndar | São Paulo, Rio | Hedonic (listing + contracts) | 2023 |
| SECOVI-SP | SECOVI-SP | São Paulo | Median prices (transactions) | 2008 |
